Automatic Parallelization and Transparent Fault Tolerance\\ (New project paper)

### 2016-04-05 pubs/tfp16/paper.tex

More generally, \emph{task-based} parallelism is becoming increasingly popular
for HPC, often in conjunction with a higher-level (typically inter-node) bulk
synchronous model typically implemented with the Message Passing Interface
(MPI) \cite{MPI}.  In the imperative world, example systems include the
aforementioned Legion, Intel's Threading Building Blocks~\cite{Reinders:2007},
and the HPX C++11/14 extension~\cite{HPX}.  While the concept of task is
somewhat vague, the general idea is a unit of computation with no or
minimal/constrained side effects.  A function in a pure functional language
could be regarded as the essence of the concept of task.  All of these
systems are of course in languages with strict function-call semantics.

@@

We perform standard Hindley-Milner type inference and
enforcement~\cite{Heren02}, augmented with built-in and user-defined unboxed
types as described by Peyton-Jones and Launchbury~\cite{Jones91unboxedvalues},
but for unboxed types code generation and runtime support is currently limited
to built-in single-word (64-bit) or smaller types (e.g., \texttt{Int\#},
\texttt{UInt\#}, \texttt{Float\#}, \texttt{Double\#}) and user-defined
enumerations, e.g., \texttt{data unboxed Bool\# = False\# | True\#}.

@@

Currently we are developing the parallel runtime.  This summer (2016) a
returning student will (partially) bridge the gap between GHC STG (or Core)
and our STG, and augment our runtime as needed in this effort.  Just as in the
Intel Haskell Research Compiler effort, we will not attempt to reuse the GHC
runtime.

>>>>

More generally, \emph{task-based} parallelism is becoming increasingly popular
for HPC, often in conjunction with a higher-level (typically inter-node) bulk
synchronous model typically implemented with the Message Passing Interface
(MPI) \cite{MPI}.  In the imperative world, example systems include the
aforementioned Legion, Intel's Threading Building Blocks~\cite{Reinders:2007},
and the HPX C++11/14 extension~\cite{HPX}.  While the concept of task is
somewhat vague, the general idea is a unit of computation with no or
minimal/constrained side effects.  A function in a pure functional language
could be regarded as the essence of the concept of a task.  All of these
systems are of course in languages with strict function-call semantics.

@@

We perform standard Hindley-Milner type inference and
enforcement~\cite{Heren02}, augmented with built-in and user-defined unboxed
types as described by Peyton-Jones and Launchbury~\cite{Jones:1991},
but for unboxed types code generation and runtime support is currently limited
to built-in single-word (64-bit) or smaller types (e.g., \texttt{Int\#},
\texttt{UInt\#}, \texttt{Float\#}, \texttt{Double\#}) and user-defined
enumerations, e.g., \texttt{data unboxed Bool\# = False\# | True\#}.

@@

Currently we are developing the parallel runtime.  This summer (2016) a
returning student will bridge the gap between GHC STG (or Core) and our STG,
and augment our runtime as needed in this effort.  Just as in the Intel
Haskell Research Compiler effort, we will not attempt to reuse the GHC
runtime.
